Add Plant Functional Type (PFT) labels

### 2021-03-29 R/add_plant_functional_type.Rmd
starting to add PFT labels manually 7/146 done lol

---
title: "Add Plant Functional Type labels"
output: html_document
---

@@

## USDA Plants Database

@@

```{r}
#usda_plants <- read_csv(file.path(here::here(), "analysis", "data", "usda_plants_20160713.csv"))

@@

## BIEN

@@

```{r install_load_packages_bien, message=FALSE}
library(BIEN)
library(ape) # Package for working with phylogenies in R
library(maps) # Useful for making quick maps of occurrences
library(sp) # A package for spatial data
```

"Let’s say we’re interested in the species Xanthium strumarium and we’d like some occurrence data."

```{r}
Xanthium_strumarium <- BIEN_occurrence_species(species = "Xanthium strumarium")
#View(Xanthium_strumarium)
head(Xanthium_strumarium)
str(Xanthium_strumarium)
```

@@

#View(Abies_taxonomy)
Abies_taxonomy[1:10,]
```

*Acer* genus is maple trees (deciduous trees). The "higher_plant_group" returned by the BIEN taxonomy is "flowering plants".

```{r}
Acer_taxonomy<-BIEN_taxonomy_genus(genus = "Acer")

#View(Acer_taxonomy)
Acer_taxonomy[1:10,]
```
*Acer rubrum* species is the red maple (deciduous trees). The "higher_plant_group" returned by the BIEN taxonomy is "flowering plants".

```{r}
Acer_rubrum_taxonomy <- BIEN_taxonomy_species(species = "Acer rubrum")
Acer_rubrum_taxonomy[1:10,]
```
```{r}
Artemisia_tridentata_taxonomy <- BIEN_taxonomy_species(species = "Artemisia tridentata")
Artemisia_tridentata_taxonomy[1:10,]
```
```{r}
Carpinus_caroliniana_taxonomy <- BIEN_taxonomy_species(species = "Carpinus caroliniana")
Carpinus_caroliniana_taxonomy[1:5,]
```

```{r}
Juniperus_virginiana_taxonomy <- BIEN_taxonomy_species(species = "Juniperus virginiana")
Juniperus_virginiana_taxonomy[1:nrow(Juniperus_virginiana_taxonomy),]
```
```{r}
Quercus_laevis_taxonomy <- BIEN_taxonomy_species(species = "Quercus laevis")
Quercus_laevis_taxonomy[1:nrow(Quercus_laevis_taxonomy),]
```

## TRY Database

@@

## Taxize

@@

>>>>

---
title: "Add Plant Functional Type (PFT) labels"
output: html_document
---

@@

Plant functional types (PFTs) of interest: coniferous forest, deciduous forest, woody shrub, grass/herb, and bare ground.

## R packages

### USDA Plants Database

@@

```{r, include = FALSE, eval=FALSE}
#usda_plants <- read_csv(file.path(here::here(), "analysis", "data", "usda_plants_20160713.csv"))

@@

### BIEN

@@

```{r install_load_packages_bien, message=FALSE, warning=FALSE}
library(BIEN)
library(ape) # Package for working with phylogenies in R
library(maps) # Useful for making quick maps of occurrences
library(sp) # A package for spatial data
```

@@

#View(Abies_taxonomy)
Abies_taxonomy[1:3,]
```

*Acer rubrum* species is the red maple (deciduous trees). The "higher_plant_group" returned by the BIEN taxonomy is "flowering plants".

```{r}
Acer_rubrum_taxonomy <- BIEN_taxonomy_species(species = "Acer rubrum")
Acer_rubrum_taxonomy[1:3,]
```

### TRY Database

@@

### Taxize

@@

## Manually add PFTs to spectral library

```{r load_cleaned_spectra, message=FALSE}
# read cleaned spectral library file 
cleaned_spectra <- read_csv(here::here("analysis", "data", "derived_data", "cleaned_spectra.csv"))
```

```{r}
# list unique taxon labels, add an empty columns to fill with plant functional type (PFT) and common name labels.
# PFTs of interest: coniferous_forest, deciduous_forest, woody_shrub, grass_herb, and bare_ground.
taxon_pft_lookup <- data.frame(genusSpecies = unique(cleaned_spectra$genusSpecies),
                               commonName = NA,
                               pft = NA)

for(i in 1:nrow(taxon_pft_lookup)){
  # print row and current genusSpecies label
  print(i)
  genusSpecies <- taxon_pft_lookup$genusSpecies[i]
  print(genusSpecies)

  if(genusSpecies == "Pseudotsuga menziesii"){
    taxon_pft_lookup$pft[i] = "coniferous_forest"
    taxon_pft_lookup$commonName[i] = "Douglas-fir"}

  else if(genusSpecies == "Alnus rubra"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"  # deciduous broadleaf tree 
    taxon_pft_lookup$commonName[i] = "red alder"}

  else if(genusSpecies == "Vaccinium parvifolium"){ 
    taxon_pft_lookup$pft[i] = "woody_shrub"       # deciduous Shrub
    taxon_pft_lookup$commonName[i] = "red huckleberry"}

  else if(genusSpecies == "Mahonia nervosa"){
    taxon_pft_lookup$pft[i] = "woody_shrub"      # evergreen shrub
    taxon_pft_lookup$commonName[i] = "Cascade barberry"}

  else if(genusSpecies == "Corylus cornuta"){
    taxon_pft_lookup$pft[i] = "woody_shrub"      # deciduous shrubby hazel
    taxon_pft_lookup$commonName[i] = "beaked hazelnut"}

  else if(genusSpecies == "Gaultheria shallon"){
    taxon_pft_lookup$pft[i] = "woody_shrub"     # broadleaved evergreen shrub
    taxon_pft_lookup$commonName[i] = "salal"}

  else if(genusSpecies == "Acer circinatum"){
    taxon_pft_lookup$pft[i] = "deciduous_forest"
    taxon_pft_lookup$commonName[i] = "vine maple"}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else if(genusSpecies == ""){
    taxon_pft_lookup$pft[i] = ""
    taxon_pft_lookup$commonName[i] = ""}

  else print("PFT unknown")

  # print corresponding PFT
  print(taxon_pft_lookup$pft[i])
}

taxon_pft_lookup$pft(taxon_pft_lookup$genusSpecies == "Pseudotsuga menziesii") <- "coniferous_forest"
```
